---
output:
  html_document:
    df_print: paged
  pdf_document: default
runtime: shiny
title: "Lab1"
---
# Lab1-Q2
```{r load libraries, echo=FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(dplyr)
library(gridExtra)
library(grid)
library(plotly)
library(shiny)
```

## Lab1-Q2-1
Read data from SENIC.txt into R.

```{r clean_data, echo=FALSE}
# loading file, changing column name
load_file <- function(file_name){
  df <- read.table(file_name, sep = "",header = FALSE)
  colnames(df) <- c("ID", "X1", "X2", "X3", "X4", "X5", "X6", "X7", "X8", "X9", 
                    "X10", "X11")
  return (df)
}
```
```{r clean_data_loading, echo=TRUE}
data = load_file('material/SENIC.txt')
head(data)
```
## Lab1-Q2-2
Create a function that for a given column (vector) X does the following:
a. It computes first and third quartiles Q1 and Q3 with quantiles()
b. It returns indices of outlying observations, i.e. observation with X-values 
greater than Q3+1.5(Q3-Q1) or less than Q1-1.5(Q3-Q1).

```{r computational_function, echo=TRUE}

get_observations_according_to_quartile <-function(data, col_name){
  # create quartiles
  quartiles = quantile(unlist(data[col_name]), 
                       prob=c(.25, .5, .75, 1))
  # naming quarties
  names(quartiles) <- c("Q1", "Q2", "Q3", "Q4")
  # Q3+1.5(Q3-Q1)
  formula_1 = quartiles[['Q3']]+ 1.5 * (quartiles[['Q3']]-quartiles[['Q1']])
  # Q1-1.5(Q3-Q1)
  formula_2 = quartiles[['Q1']]-1.5 * (quartiles[['Q3']]-quartiles[['Q1']])
  return (data %>%
           mutate('outlier' = ifelse(
             (!!sym(col_name) > formula_1 | !!sym(col_name) 
                                 < formula_2), !!sym(col_name), NA))
          )
}
create_density_plot <- function(data, outlier_data, 
                                column_to_create_density_graph, x_label){
  if (nrow(outlier_data) >=1){
    # if outlier data is available
    p <- ggplot(data, aes_string(x=column_to_create_density_graph)) + 
      geom_density() +
      xlab(x_label) +
      geom_point(data=outlier_data, 
                 aes(x=outlier, y=0), shape=18)
  }
  else{
    # if outlier data is not available
    p <- ggplot(data, aes_string(x=column_to_create_density_graph)) + 
      geom_density() +
      xlab(x_label)
    
  }
  return (p)
}
create_density_plot_with_bw <- function(data, outlier_data, 
                                column_to_create_density_graph, x_label, bw){
  if (nrow(outlier_data) >=1){
    # if outlier data is available
    p <- ggplot(data, aes_string(x=column_to_create_density_graph)) + 
      geom_density(bw=bw) +
      xlab(x_label) +
      geom_point(data=outlier_data, 
                 aes(x=outlier, y=0), shape=18)
  }
  else{
    # if outlier data is not available
    p <- ggplot(data, aes_string(x=column_to_create_density_graph)) + 
      geom_density(bw=bw) +
      xlab(x_label)
    
  }
  return (p)
}

outlier_data_and_graph <- function(column_name, x_label){
  outlier_data <- get_observations_according_to_quartile(data, column_name)  
  outlier_data <- outlier_data %>%
                            filter(!is.na(outlier))
  graph <- create_density_plot(data, outlier_data, column_name, x_label)

    
  return (graph)
}

outlier_data_and_graph_with_bw <- function(column_name, x_label, bw){
  outlier_data <- get_observations_according_to_quartile(data, column_name)  
  outlier_data <- outlier_data %>%
                            filter(!is.na(outlier))
  graph <- create_density_plot_with_bw(data, outlier_data, column_name, x_label,
                                       bw)

    
  return (graph)
}
```
## Lab1-Q2-3

Use ggplot2 and the function from step 2 to create a density plot of Infection 
risk in which outliers are plotted as a diamond symbol diamond . Make some 
analysis of this graph.

```{r graph_function, echo=FALSE}
# as X3 is infection risk
infection_risk.graph = outlier_data_and_graph("X3", "Infection Risk")
print(infection_risk.graph)
```


#### comments:
With the density graph it is clearly visible that average of estimated probility
of acquiring infection in hospital is analogous to normal distribution. More 
hospitals are at average infection risk. Whereas outliers tells that very few 
are less or zero risk or at very high risk.

## Lab1-Q2-4
Produce graphs of the same kind as in step 3 but for all other quantitative 
variables in the data (aes_string() can be useful here). Put these graphs into
one (hint: arrangeGrob() in gridExtra package can be used) and make some 
analysis.

```{r all_graphs_arrange_function, echo=FALSE}
#### length of stay
# filtering outlier_data
# Length of Stay = X1
length_of_stay.graph <- outlier_data_and_graph("X1", "Length of Stay")
# Age
age.graph <- outlier_data_and_graph("X2", "Age")
# routine culturing ratio
routine_culturing_ratio.graph <- outlier_data_and_graph("X4", 
                                                      "Routine Culturing Ratio")
# routine chest x-ray ratio
routine_chest_x_ray_ratio.graph <- outlier_data_and_graph("X5", 
                                                    "Routine Chest X-Ray Ratio")
# number of beds
number_of_beds.graph <- outlier_data_and_graph("X6", "Number of Beds")
# average daily census
average_daily_census.graph <- outlier_data_and_graph("X9", 
                                                     "Average Daily Census")
# number of nurses
number_of_nurses.graph <- outlier_data_and_graph("X10", "Number of Nurses")
# available facilities and services
available_facilities_and_services.graph = outlier_data_and_graph("X11", 
                                            "Available Facilities and Services")

grid.arrange(
  length_of_stay.graph, 
  age.graph, 
  infection_risk.graph, 
  routine_culturing_ratio.graph,
  routine_chest_x_ray_ratio.graph,
  number_of_beds.graph,
  average_daily_census.graph,
  number_of_nurses.graph,
  available_facilities_and_services.graph,
  ncol=3)
```

#### comments:


Lengh of stay: Usually length of stay from 1 to 12 days.

Age: More cases are from age after 40 and between 60. It also shows that after 
60, there are few people going to hospital. This also depicts the average of 
pateints.

Routine Culturing ratio or Chest-X-ray ratio depicts that in most hospitals chest X-ray
occurance is more than routine Culturing ratio.
Graph also depicts that most hospitals offer these facilities to normal pateints.

Number of Beds: More hospitals have average 100 beds, 
whereas hospitals with more beds are less.

Number of Nurses: Same as number of beds, more hospitals have 100 number of 
nurses whereas hopstials with more number of nurses are less.


## Lab1-Q2-5
Create a ggplot2 scatter plot showing the dependence of Infection risk on the 
Number of Nurses where the points are colored by Number of Beds. Is there any 
interesting information in this plot that was not visible in the plots in 
step 4? What do you think is a possible danger of having such a color scale?






```{r scatter_plot, echo=FALSE}
number_of_nurses_and_infection_risk.graph <- ggplot(data, aes(x=X10, y=X3, 
                                                              color=X6)) +
  geom_point() +
    xlab('Number of Nurses') +
    ylab('Infection Risk')
print(number_of_nurses_and_infection_risk.graph)
```

#### comments

We can clearly see that more the number of nurses and lesser the risk of 
infection. The color scales also depict that risk of infection is also dependent
on the number of beds. But from 0 to 200, color difference is so minor that we 
not differentiate between different number of beds.


## Lab1-Q2-6
Convert graph from step 3 to Plotly with ggplotly function. What important new 
functionality have you obtained compared to the graph from step 3? Make some 
additional analysis of the new graph.
```{r plotly_for_step3, echo=FALSE}
ggplotly(infection_risk.graph)

```


#### comments
It is same as older graph except its interactivity, which is very helpful, for example
by hovering the mouse, we can see the outlier values and can the clear difference between them.

## Lab1-Q2-7

Use data plot-pipeline and the pipeline operator to make a histogram of 
Infection risk in which outliers are plotted as a diamond symbol. 
Make this plot in the Plotly directly 
(i.e. without using ggplot2 functionality). Hint: select(), filter() and 
is.element() functions might be useful here.

```{r complete_plotly, echo=FALSE}
outlier_data <- get_observations_according_to_quartile(data, "X3")
plotly.graph <- data %>% 
                select(X3) %>% 
                plot_ly(x = ~X3, type = "histogram") %>% 
                add_markers(x = outlier_data$outlier, y = 0, showlegend = FALSE,
                marker = list(symbol = c("diamond"))) %>% 
  layout(
    title = "Histogram of Infection risk in with outliers are plotted as a \
    diamond symbol")
plotly.graph
```
## Lab1-Q2-8
Write a Shiny app that produces the same kind of plot as in step 4 but in 
addition include:
a. Checkboxes indicating for which variables density plots should be produced
b. A slider changing the bandwidth parameter in the density estimation ('bw'
parameter)
```{r shinye_app, echo=FALSE}
ui <- fluidPage(
  titlePanel("Shiny App"),
  sidebarLayout(
    sidebarPanel(
        sliderInput(
            "bwp",
            "bandwidth parameter:",
            min = 0.1, max = 10, value = 0.5
        ),

        checkboxGroupInput("cb", "Variables to show:",
                     c("Length of Stay" = "X1",
                       "Age" = "X2",
                       "Infection Risk" = "X3",
                       "Routine Culturing Ratio" = "X4",
                       "Routine Chest X-ray Ratio" = "X5",
                       "Number of Beds" = "X6",
                       "Average Daily Census" = "X9",
                       "Number of Nurses" = "X10",
                       "Available Facilities & Services" = "X11"
                       )
                     ,selected = c("X1")
                     ),
    ),

    # Main panel for displaying outputs ----
    mainPanel(
      plotOutput("outputPlot")
    )
  )
)

# Define server logic to plot various variables against mpg ----
server <- function(input, output) {
    getPlots <- function(cols, bw) {
        des_list <- c("Length of Stay", "Age", "Infection Risk", 
                      "Routine Culturing Ratio", "Routine Chest X-Ray Ratio", 
                      "Number of Beds", "Average Daily Census", 
                      "Number of Nurses",
                      "Available Facilities and Services")
        nplot<-length(cols)
        x<-cols
        p<-list()
        for ( i in 1:nplot) {
            p[[i]] <- outlier_data_and_graph_with_bw(cols[i], des_list[i], bw)
        }
        return(p)
    }
    output$outputPlot <- renderPlot({
        p <- getPlots(input$cb, input$bwp)
        do.call(grid.arrange,p)
  })

}

shinyApp(ui, server)
```

